### 2014-12-02 tex/short_2015/body.tex
more stuff

@@

We compute $\sv_{dx}$ and $\sv_{dy}$ by sampling proxy model positions and
performing the perspective division and transformation into window coordinates
in the vertex shader.
Let
\[
  \text{pos} =
  U { \small \begin{pmatrix} x \\ y \\ d \\ 1 \end{pmatrix} },
  \text{pos}_{\Delta x} =
  U { \small \begin{pmatrix} x+\Delta x \\ y \\ d_{\Delta x} \\ 1 \end{pmatrix} }
  \text{and}\, \, \, 
  \text{pos}_{\Delta y} =
  U { \small \begin{pmatrix} x \\ y+\Delta y \\ d_{\Delta y} \\ 1 \end{pmatrix} },
\]
where $d = 2\text{depthTex}(x, y) - 1$, $d_{\Delta x} =
2\text{depthTex}(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2\text{depthTex}(x,
y+\Delta y) - 1$ are depths sampled from the received depth buffer and transformed
to $[-1, 1]$.

@@

Then, for the 2D vectors spanning the splats in client window coordinates, we
get
\[
  \sv_{\Delta x} =
    \frac{\text{viewport width}}{2\Delta x} \left(
        \frac{\text{pos}_{\Delta x}.xy}{\text{pos}_{\Delta x}.w} -
        \frac{\text{pos}.xy}{\text{pos}.w}
    \right),
\]
and
\[
  \sv_{\Delta y} =
    \frac{\text{viewport width}}{2\Delta y} \left(
        \frac{\text{pos}_{\Delta y}.xy}{\text{pos}_{\Delta y}.w} -
        \frac{\text{pos}.xy}{\text{pos}.w}
    \right),
\]
where we have used the ``shader notation'' to extract 2D vectors and scalars
from the 4D vectors.

@@

@@

@@

In the case of splat sets: Replacement algorithms

@@

@@

@@

%-------------------------------------------------------------------------
\subsection{References}

List all bibliographical references in 9-point Times, single-spaced, at the
end of your paper in alphabetical order. When referenced in the text, enclose
the citation index in square brackets, for example~\cite{Lous90}. Where
appropriate, include the name(s) of editors of referenced books.

For your references please use the following algorithm:
\begin{itemize} 
\item \textbf{one} author: first 3 chars plus year -- 
      e.g.\ \cite{Lous90}
\item \textbf{two}, \textbf{three} or \textbf{four} authors: first char
      of each family name plus year --  e.g.\ \cite{Fellner-Helmberg93} 
      or \cite{Kobbelt97-USHDR} or \cite{Lafortune97-NARF}
\item \textbf{more than 4} authors: first char of family name from 
      first 3 authors followed by a '*' followed by the year -- 
      e.g.\ \cite{Buhmann:1998:DCQ} or \cite{FolDamFeiHug.etal93} 
\end{itemize}

For BibTeX users a style file \ \texttt{eg-alpha.bst} \ is available which
uses the above algorithm.

%-------------------------------------------------------------------------
\subsection{Conclusions}

Please direct any questions to the production editor in charge of
these proceedings.

>>>>

%-------------------------------------------------------------------------
\section{Notes}

\begin{itemize}

\item We restrict ourselves to discuss ``splat sets'' in this paper. (Nevne der det
tas frem andre alternativer)

\item
Husk aa faa med det viktige i at vi faar oekt serverkapasiteten til aa
serve flere klienter ogsaa!

\item
``server-image'' i steden for ``server-rendered image'' for aa spare plass?

\item received stavekontroll

\item Hvor passer det aa faa inn 1) bilde og 2) tekst om FRView?

\item Maa ta et lite soek paa prior art innen omraadet

\end{itemize}

@@

% After mult by proj matrix: clip coo, range of (x, y, z) = [-w, w]^3
% After persp division: ndc [-1, 1]^3
% After viewport transf: window coordinates
% (Range [vp[0] vp[1]] x [[vp[2] vp[3]] x [0, 1] ?)

We compute $\sv_{dx}$ and $\sv_{dy}$ by evaluating proxy model positions and
performing the perspective division and transformation into window coordinates
in the vertex shader. Let proxy model positions corresponding to three corners
of a quadrilateral splat in clip coordinates be
given by
\[
  \pv =
  U { \small \begin{pmatrix} x \\ y \\ d \\ 1 \end{pmatrix} },
  \pv_{x+\Delta x} =
  U { \small \begin{pmatrix} x+\Delta x \\ y \\ d_{\Delta x} \\ 1 \end{pmatrix} }
  \text{and}\, \, \, 
  \pv_{y+\Delta y} =
  U { \small \begin{pmatrix} x \\ y+\Delta y \\ d_{\Delta y} \\ 1 \end{pmatrix} },
\]
where $d = 2\text{depthTex}(x, y) - 1$, $d_{\Delta x} =
2\text{depthTex}(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2\text{depthTex}(x,
y+\Delta y) - 1$ are depths sampled from the received depth buffer and
transformed to $[-1, 1]$. By combining this with a $w$-component set to one, we
have in effect done a perspective division, so that the right-hand side vertices
are in clip space, suitable for multiplication with $U$.

@@

\textbf{ Litt om valgene vi kan gjoere for Delta}

This again leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
coordinates, and after perspective division and viewport transformation, we get
corresponding window coordinates $\sv$, $\sv_{x+\Delta x}$ and $\sv_{y+\Delta
y}$, which we can subtract to obtain 2D vectors spanning the splat,
\[
  \sv_{\Delta x} =
  \sv_{x+\Delta x} - \sv =
    \frac{\text{width}}{2\Delta x} \left(
        \frac{\pv_{x+\Delta x}.xy}{\pv_{x+\Delta x}.w} -
        \frac{\pv.xy}{\pv.w}
    \right),
\]
and
\[
  \sv_{\Delta y} =
  \sv_{y+\Delta y} - \sv =
    \frac{\text{height}}{2\Delta y} \left(
        \frac{\pv_{y+\Delta y}.xy}{\pv_{y+\Delta y}.w} -
        \frac{\pv.xy}{\pv.w}
    \right),
\]
where we have used the ``shader notation'' to extract 2D vectors and scalars
from the 4D vectors, and $\text{width}$ and $\text{height}$ are the viewport
width and height, respectively.

@@

The problem with this, is that the geometry is not planar, and our simple texture
coordinate transformation will produce the wrong result. Some ways to remedy
this, is to choose smaller (and more) splats, detect the problem and discard
such splats, or introduce more complex texture transformations. Since the proxy
geometry in principle is to be short-lived, and only a temporary geometry to be
rendered before the server-rendered image can be receieved, we have chosen the
first alternative. Discarding such splats may offer a good alternative,
especially when combining several proxy geometries, where the union of retained
splats increases the likelihood of all geometry parts being covered. Note also
that implementing a more sophisticated texture coordinate transformation may
actually amount to performing the same work as for more and smaller splats. Or,
in other words, using more and smaller splats may be regarded as a better
texture transform implementation.

@@

For larger splats, covering many fragments, it makes sense to also compute and
use depth fragments. This will improve the usage of more proxy models
simultaneously, since the risk of a splat with a not very representative depth
value covers another and better splat is reduced. The ``intra-splat'' depth
values can easily be fetched from the depth image, just as the texture is looked
up for color. One problem is that many thin clients may not support this, in
WebGL, the ``frag depth'' feature is an extension, which may not be supported
everywhere.

@@

In this paper we concern ourselves with proxy models defined as sets of
splats. Since each splat set is the product of one instance of a server-rendered
image, depth buffer and viewing parameters, it makes sense to retain more than
one such model on the client. By doing this, the client may combine them, in
order to provide renderings for viewing parameters very different from the last
one received. One can imagine a plethora of splat set replacement algorithms, we
have tested two approaches which both works well; both retain a constant number
of proxy models. The first algorithm is to replace the one with a viewing
direction differing the most from the newly received model. The second simply
replaces the oldest one in store. The third replaces a proxy model $j$ if the
replacement results in the following objective function being reduced,
\[
  \text{coverage}_j = 
  \sum_{i=1, i\neq j}^n 
    \angle(\textbf{camdir}_i, \textbf{new camdir} )^2,
\]
where $\textbf{camdir}$ is the direction in which the camera was looking for the
generation of that particular proxy model.

There are some issues to consider with respect to panning and zooming that we
will not discuss here due to the limited space available.  All of these model
replacement algorithms work decently well.

@@

Since the main purpose of the automatically generated proxy model is to
facilitate a client-rendered image while waiting for one from the server for
increased interactivity, it is important that the process of generating and
sending proxy model data from the server itself is not hold up more than
necessary. There are mainly three sources of delay (\textbf{passer aa nevne
tidligere?}) for server-rendered images to the client; high
latency, low bandwidth, and slow server-rendering itself.

In the first case, it seems prudent to have a better proxy model on the client,
that can be used for longer time and for interactivity resulting in viewing
parameters deviating more from the last received server-rendered image. In the
two other cases, it is important for the proxy model generation/transmission to
be quicker, both in order to get the proxy model to the client and keep from
delaying the server-image more than necessary.

To take care of this, we have adopted an adaptive specification of
proxy model data from the server. This involves a more light-weight image (lossy
JPG-compression with adaptive quality control) while the user is interacting
with the client, as well as a reduced resolution of the depth buffer sent from
the server. See Figure~\ref{fig:adapativeProxyModels} for examples of low- and
high-quality proxy models.

\begin{figure}[htb]
  \centering
  \subfigure[Low quality (tenker her jpg med q=0 og daarlig dybdebuffer, faa
  splats kanskje]{
    \input{fig3a.tex}
  }
  \subfigure[High quality (png med maks av alt, mange splats)]{
    \input{fig3b.tex}
  }
  \caption{\label{fig:adapativeProxyModels}
           Proxy models adaptively adjusted to available bandwidth and server
           capacity.}
\end{figure}

@@

We have tested the algorithms discussed in the Tinia framework, which is a
programming framework for setting up and managing a client/server based
interactive OpenGL-based visualization application. The client is
Javascript/WebGL-based and runs in one of several possible browsers, \eg Chrome
or Firefox. This client communicates over tcp/ip, using either http or
websockets, with a server running as an Apache module.

The automatic proxy geometry implementation is fully invisible to the
application, meaning that all existing applications immediately will have the
feature available. We have tested the algorithm on several smaller test cases,
but also on a larger oild reservoir viewer.

This viewer utilizes several GLSL shaders to render reservoir cells and
boundaries, tubular wells etc., and operates in general on a geometric model
that is not trivial to reduce in complexity. The model is typically also very
large, meaning that rendering on a thin client is prohibitive. With the
automatic proxy geometry, we can obtain interactive frame rates with a bad
connection from a very lightweight client to the server. For a comparison of a
server-rendered image and a client-rendered proxy model, see
Figure~\ref{fig:FRView}.

\begin{figure}[htb]
  \centering
  \subfigure[Server-rendered]{
    \input{fig3a.tex}
  }
  \subfigure[Client-rendered proxy model]{
    \input{fig3b.tex}
  }
  \caption{\label{fig:FRView}
           The oil reservoir viewer FRView}
\end{figure}

@@

The single most attractive feature of such an automatic proxy geometry as
described, is exactly the automated generation. It means that problems with
downscaling, sending and rendering existing geometries is bypassed
altogether. The Tinia framework with which we have demonstrated this, is a
framework for setting up and managing a client/server based interactive OpenGL
based visualization application. The automatic proxy geometry implementation is
fully invisible to the application, meaning that all existing applications
immediately will have the feature available.

There are several directions in which we would like to follow up and improve
this concept,
\begin{itemize}
\item \textbf{Proxy model replacement algorithms} The goal is to obtain a good view with
all relevant viewing parameters on the client, with the help of a minimal set of
proxy models.
\item \textbf{Grid-based proxy geometry} Instead of a set of ``splats'', one can
envision a spatially partioned grid, for instance containing a distance field,
with the geometry as an iso-surface.
\end{itemize}
